Triangulating iconicity: Coding, consistency & calibration

### 2021-06-09 TI_03_ratings.Rmd
cleaning up and refactoring for release

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

@@

>>>>

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis","gghalves","plotly","lme4","VGAM","plyr","emmeans","car")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

@@

## Rating: descriptives

```{r}
# load dataframe with all ratings
read.csv("data\\ideophones_rated_uncorr.csv",header=TRUE,na.strings=c("", "NA"),sep=";",check.names=FALSE) -> d.uncorr

# to long format
d.uncorr <- gather(d.uncorr, item, rating, 6:245)
d.uncorr <- d.uncorr[,c(1,6,7)]
colnames(d.uncorr) <- c("rater","item","rating")
#remove empty cells
d.uncorr <- subset(d.uncorr, !is.na(d.uncorr[,"rating"]))
#change ratings to numeric values with decimals
as.numeric(sub(",", ".", d.uncorr$rating, fixed = TRUE)) -> d.uncorr$rating

# check for inconsistent raters with person-total correlation based on Motamedi et al. (2019)
ptc <- d.uncorr[,c("item","rater","rating")]
personTotalCorrelationCorrected <- function(ptc)
  {
  require(tidyverse)
  ptc %>%
    dplyr::rename(raterFocal = rater,
           ratingFocal = rating) %>%
    full_join(ptc, by='item') %>%
    filter(rater!=raterFocal) %>%
    group_by(item, raterFocal, ratingFocal) %>%
    summarise(ratingsOthers = mean(rating, na.rm=TRUE)) %>%
    base::split(.$raterFocal) %>%
    map(~cor(.$ratingFocal, .$ratingsOthers, 
             use="pairwise.complete.obs")) %>%
    as.data.frame %>%
    gather %>%
    dplyr::rename(rater = key, perTotCor = value)
}
View(personTotalCorrelationCorrected(ptc))

#plot person-total correlations
ptc %>%
select(item, rater, rating) %>%
personTotalCorrelationCorrected %>%
ggplot(aes(x=perTotCor)) + 
  geom_density() + 
  geom_histogram(stat="bin", 
                 alpha=0.4,
                 aes(y=..density..),
                 binwidth=0.1) +
  labs(x="Person-total correlation")
ggsave("figures\\rating_PerTotCorr.png",height=5,width=7.5)
```

One participant (N045) showed a negative person-total correlation, which according to Curran (2016) indicates a careless responder rendering invalid data. Inspection of this participant’s data showed they had indeed given the same answer to every question. I therefore decided to remove this participant from the dataset.

Two other participatns (N022 and N034) also had to be removed because of their knowledge of Japanese and Korean, resulting in a total of 75 participants to be included in the analysis.

```{r}
# load corrected data
d <- read_xlsx("data\\ideophones_rated.xlsx")
d.means <- read_xlsx("data\\ideophones_rated_means.xlsx")

# get means and standard deviations across all ideophones and for each language and category
mean(d$rating)
  # 2.948339
sd(d$rating)
  # 1.297106
sumcat <- ddply(d,~category,summarise,mean=mean(rating),sd=sd(rating))
sumlang <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))

Because the category "Other" is quite a vague residual category with only 17 items (as opposed to 30-40 items for the other categories), I decided to remove this category for now. I will keep a copy that includes the ratings for "Other".

```{r}
#create copy of data including the category "Other"
d.other <- read_xlsx("data\\ideophones_rated.xlsx")
d.m.other <- read_xlsx("data\\ideophones_rated_means.xlsx")

#remove category "Other" from d and d.means
d <- d[!grepl("Other",d$category),]
d.means <- d.means[!grepl("Other",d.means$category),]

# make new summary for language (without the Japanese "other" ideophones)
sumlang2 <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))
```

Here we plot the distribution of the rating data.

```{r}
#convert some columns to factors
cols <- c("category","language","study")
d[cols] <- lapply(d[cols], factor)
d.means[cols] <- lapply(d.means[cols], factor)

#combined boxplot and scatterplot by category
ggplot(d.means, aes(x=reorder(category,-rating),y=rating,color=category)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
  "Texture","Colour/Visual"))
ggsave("figures\\rating_category.png",height=5,width=7.5)

#same, but by language
ggplot(d.means, aes(x=reorder(language,-rating),y=rating,color=language)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category")
ggsave("figures\\rating_language.png",height=5,width=7.5)

## Statistics

Now we want to see which categories and languages differ significantly from the others in their iconicity ratings.

First, we check the assumptions.

```{r}
# check normality by visual inspection
ggqqplot(d, "rating", facet.by = "category")
  # all points fall approximately along the reference line, so normality can be assumed

# check homogeneity of variance with Levene's test
leveneTest(rating~category,data=d)
leveneTest(rating~language,data=d)
  # both are not significant (> .05), so homogeneity of variance can be assumed
```

Now we construct a linear model with category as a predictor of the iconicity ratings.

```{r}
summary(lm.cat <- lm(rating ~ category, d))

# pairwise comparisons with Tukey adjustments in the emmeans package
emm.cat <- emmeans(lm.cat, "category")
pairs(emm.cat)
pwpm(emm.cat)
```

And here we do the same with language as predictor.

```{r}
summary(lm.lang <- lm(rating ~ language, d))

# pairwise comparisons in emmeans package
emm.lang <- emmeans(lm.lang, "language")
pairs(emm.lang)
pwpm(emm.lang)
```

## Combined data
Here we load the combined coding and guessability data and then add the rating data.

```{r}
# get coded and guessed data from previous step (TI_02_congruence)
d = read.csv("data\\ideophones_coded_guessed.csv")

coding_categories <- d %>% dplyr::select(matches('F_|M_'),-matches('notes|meaning')) %>% names()

# add mean ratings
d.ratings = read_excel("data\\ideophones_rated_means.xlsx") %>%
  dplyr::select(-c(item,category,list))
d <- left_join(d,d.ratings,by=c("filename","language","study"))

# convert ratings to z-scores for comparison with guessability (logodd) scores
d$rating_z <- scale(d$rating,center=T,scale=T)

#remove category "Other"
d <- d[!grepl("Other",d$category),]

## Visualisations
Let's first visualise the relation between the rating and guessing data.

```{r}
ggplot(d,aes(x=rating_z,y=score_z)) +   
  geom_point(aes(colour=study),position="jitter") +
  geom_smooth(method=lm,colour="black") +
  theme_tufte(base_size = 16) +
  xlab("rating") + ylab("guessability")
ggsave("figures\\rating_guessing_corr.png",height=5,width=7.5)

ggplot(d,aes(x=rating_z,y=score_z)) +   
  geom_point(aes(colour=category,shape=study),position="jitter") +
  geom_smooth(method=loess,colour="black") +
  theme_tufte(base_size = 16) +
  xlab("rating") + ylab("guessability")
```

There appears to be a positive correlation between the rating and guessing data.

Now let's visualise the relations between the coding, guessing and rating data.

```{r}

ggplot(d,aes(x=rating_z,y=logodds, color=C_cumulative)) +
  geom_point() +
  geom_smooth(method=loess) +
  scale_fill_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
  NULL

plot_ly(data=d,x=rating_z,y=score_z,color=C_simple)

## Statistics
```{r}
# Pearson correlation
cor.test(d$score_z,d$rating_z,method="pearson")
cor.rank <- cor.test(d$score_z,d$C_cumulative,method="spearman")
```

Pearson's correlation coefficient indicates that there is indeed a strong positive correlation between the guessing and rating data, r(220) = .61, p < .001.

```{r}
#check where the guessing and rating data deviate
d %>%
  filter(rating_z-score_z > 2 | rating_z-score_z < -2)%>%
  dplyr::select(filename,ideophone,rating,rating_z,language,logodds,score,score_z) %>%
  arrange(score) %>% ungroup()
```

### 2021-06-09 paper_coding_1_consistency.Rmd
cleaning up and refactoring for release

F_consistency$agree <- NA
F_consistency$kappa <- NA
for (i in 1:(length(F_categories)-1)) { # -1 to exclude the notes column
  F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
  F_consistency$kappa[i] <- (get_kappa(data=d.F,cat=F_categories[i]))
}
F_consistency
```

@@

For Meaning, mean agreement in the original coding is **`r round(mean.na(M_consistency$agree))`%** and the lowest Kappa values are for M-durative, M_weight, and M_irregular, pointing to the need to resolve inconsistencies and sharpen those coding categories.

@@

Hmm, a consistency check at this point shows we're nearly there but we're not 100% consistent on F_redup, F_monosyllabic and F_vowelquality and F_finallength. We can export only those columns to have a quick check.

@@

* Ultimately we want to output a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses.

```{r write_final}

# for now I'm taking just one coder
d.coded <- read_excel("data\\ideophones_coded_SP.xlsx") %>% 
  arrange(filename) %>%
  dplyr::select(-matches('notes'))
names(d.coded) <- gsub('_SP','',names(d.coded))

@@

write_xlsx(d.coded,path="data\\ideophones_coded.xlsx")

@@

>>>>

F_consistency$agree <- NA
F_consistency$kappa <- NA
for (i in 1:(length(F_categories)-1)) { # -1 to exclude the notes column
  F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
  F_consistency$kappa[i] <- (get_kappa(data=d.F,cat=F_categories[i]))
}
F_consistency
mean.na(F_consistency$kappa)

@@

For Meaning, mean agreement in the original coding is **`r round(mean.na(M_consistency$agree))`%** and the lowest Kappa values are for M_durative, M_weight, and M_irregular, pointing to the need to resolve inconsistencies and sharpen those coding categories.

@@

A consistency check at this point shows we're nearly there but we're not 100% consistent on F_redup, F_monosyllabic and F_vowelquality and F_finallength. We can export only those columns to have a quick check. However, as we do not aim for 100% consistency, this is more for informative purposes.

@@

### Consensus version
The final result of the coding consistency is a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses. We do this this by taking the (calibrated and revised) data from one coder.

```{r create_consensus_version}

d.coded <- read_excel("data\\ideophones_coded_SP.xlsx") %>% 
  arrange(filename) %>%
  dplyr::select(-matches('notes'))
names(d.coded) <- gsub('_SP','',names(d.coded))

@@

@@

## Adding congruency measures
We add congruency measures by combining F and M features as described in the coding scheme.

```{r congruency_1}

# note that we overwrite the old d with the new consensus version here

# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
d <- d.coded %>%
  mutate(C_modality = ifelse(M_sound == 1,1,0),
         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
         C_long = ifelse(F_finallength == 1 & M_long == 1,1,0),
         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight == 2),1,0),
         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
         )

# some mappings are mutually exclusive, so we combine them

#sanity check: are they indeed mutually exclusive? these calls should come up empty
d %>% 
  mutate(C_aspect = C_iterative + C_punctual) %>%
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_aspect) %>%
  filter(C_aspect > 1)
d %>%
  mutate(C_length = C_long + C_closure) %>%
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) %>%
  filter(C_length > 1)

d <- d %>%
  mutate(C_aspect = C_iterative + C_punctual,
         C_length = C_long + C_closure)

# C_weight_ subfeatures are linked to one and the same meaning feature M_weight so they are recombined into C_magnitude
# (note: this may be subject to revision, as the subfeatures may not deserve equal weight)

d <- d %>%
  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone,
         C_magnitude = ifelse(C_weight == 0,0,1)) # i.e. a magnitude value that is 1 if any weight feature was congruent

# add cumulative iconicity scores (and some others)
# any1 = 1 if an ideophone has at least 1 congruency feature of any type
# atleast2 = if an ideophone has at least 2 congruency features of any type
# ternary = simple way of conflating the (smaller) upper regions: 1=1, 2=2, >2 = 3
d <- d %>%
  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) %>%
  mutate(C_any1 = ifelse(C_cumulative == 0,0,1),
         C_atleast2 = ifelse(C_cumulative > 1,1,0),
         C_ternary = ifelse(C_cumulative == 0,0,
                            ifelse(C_cumulative == 1,1,
                                   ifelse(C_cumulative == 2,2,3)))) # i.e. 1, 2 or 3

# we can have a first look at the basic distribution of guessability by iconicity
d %>%
  group_by(C_cumulative) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_any1) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_atleast2) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_ternary) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_simple) %>%
  summarise(n=n(),score=mean.na(score))

# finally write this to XLS and CSV
write_xlsx(d,path="data\\ideophones_coded.xlsx")
write.csv(d,file="data\\ideophones_coded.csv")
